Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-12 WRITEUP/ms.tex
clarify sec 3.3

\begin{abstract}
We study the dynamical properties of molecular cloud complexes (MC) in a prototypical galaxy (\flower) and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
%
We identify MCs using H$_2$ density, and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$) with those observed in nearby and \z$\sim$\,2 galaxies.
%
The MCs span a mass range of $10^{6.5-9}$\,\Msun. Their velocity dispersions and gas surface densities are systematically higher than those observed in nearby galaxies, but comparable to starburst galaxies.
%
We compute the effective two-component Toomre $Q$ parameter (gas and stars) to examine regions that are 
gravitationally unstable.
%
We further perform virial analysis, which is accessible from observations, to examine regions likely 
to undergo collapse. Our analysis indicates that the MCs in the main disk of \flower are unbound, but its substructures have lower virial parameters and Toomre $Q$, which is consistent with the notion that collapsing structures occur within globally stable structures.
%
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback 
compared to \flower. Our results suggest that the more massive and bigger MCs compared to the Milky Way manifest from the 
higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA) can furnish scrutinizing data to test our findings and shed light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

\begin{figure*}[!htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.65\textwidth]{\figpath/sfh.pdf}
\caption{
    Star formation history of \flower (top right) and projected stellar mass distribution of \flower during {\it (a)} one of its accretion phases at its early stage of evolution; {\it (b)} one of its major starburst phases after a merger event; and {\it (c)} a relatively quiescent phase post-starburst.
\label{fig:SFH}}
\end{figure*}

@@

In the ``clump-finding'' process, we employ a set of different density thresholds defined based on the molecular hydrogen density of \flower taken at different cosmic times (between $z$\eq6.0\,$-$\,7.2). We note that this process is in essence similar to identifying molecular structures based on the noise levels of surface density maps observers obtain with telescopes, using molecular line tracers such as CO, CS, and HCN, as commonly adopted in observational studies (e.g., identifying ``clumps'' based on/after applying $\sigma$-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a}).

@@

\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples showing the MCs identified in the accreting phase of \flower by applying volumetric H$_2$ density cuts of $n_{\rm cut}$\eq[0.53, 6.81,19.00]\,cm$^{-3}$, which is a subset of all the $n_{\rm cut}$ adopted (see text). Color shows the H$_2$ surface density, weighted by the column density. 
\label{fig:MC}}
% \addtocounter{figure}{-1}
\end{figure*}

In \Fig{h2density}, we show an example of the volumetric H$_2$ density ($n_{\rm H2}$) distribution of \flower~for a given evolutionary stage which includes contributions within $R\sim$\,3.5\,kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of density where ``clumps/structures'' are found based on the Euler characteristic, i.e., the fourth Minkowsky functional (see \citealt{Pallottini17b}). We identify MCs by applying 10 equally-spaced volumetric H$_2$ density cuts of $\log{(n_{\rm cut}/{\rm cm^{-3})}}\in[-0.5, 1.5]$\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 19.00, 31.62]\footnote{We also vary the range of $n_{\rm cut}$ adopted and find no qualitative differences affecting the results and conclusions of this work except in the slope of the CMF (see \Sec{ncut}).} to the same stage used to plot \Fig{h2density}.

For visualization, we overplot the molecular structures identified using a subset of the $n_{\rm cut}$ (0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps in \Fig{MC} as an example. The plots shown in \Fig{MC} are weighted by the column density of the total gas content.
% equation below for bookkeeping:
% ${\rm pixel_i}$\eq$\frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}$,
% where $\rho_{\rm H2, i}$ is the H$_2$ density of pixel $i$, $w_i$ is the weighted quantity in pixel $i$ which we adopt
% to be the total gas density, and $\Delta z_i$ is the path length of pixel $i$ along the line-of-sight over which the quantities are integrated along.

@@

We show in \Fig{dist} the MC distributions in terms of their molecular gas masses, radii, and gas mass fractions, which we define as
\begin{equation}
f_{\rm gas} = \frac{M_{\rm cl}} {\left(M_{\rm cl} + M_*\right)},
\end{equation}
where $M_*$ is the stellar mass within the MC volume.
%
We note that the distributions vary for different evolutionary stage of \flower. The highest density threshold of $n_{\rm cut}$\eq31.62\,\cc yields a ``minimum'' MC mass of the order of 10$^{6.5}$\,\Msun for the densest MC.
%
We find $\sim$10$-$13 MCs with masses exceeding 10$^8$\,\Msun across all the considered evolutionary stages for each of the given $n_{\rm cut}$ (see \Sec{method}). The mass of the MCs identified ranges between $M_{\rm cl}\approx$10$^{6.5-9}$\,\Msun, consistent with those observed in \z$\sim$2 galaxies in rest-frame {\it UV and optical light} \citep{Elmegreen07a, Elmegreen09a}.

@@

% Relating Larson's law to stability and pressure (perhaps move this to other section?).
In general, the virial theorem for a distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot I = 2(T - T_{\rm ext}) + W,
\end{equation}
where $\ddot I$ is the second time derivative of the Lagrangian moment of inertia, $T$ is the internal energy of the gas, $T_s$ is external pressure support, and $W$ is the gravity energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and velocity dispersion $\sigma$, accounting for both thermal and turbulent velocity dispersion. Neglecting magnetic field and calling $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}\ddot I = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R - \frac{\Gamma GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the first term on the right hand side (RHS) is the pressure term from internal velocity dispersion, the second term the external pressure term, and the third term is is the gravity term. For the case of simple virial equilibrium (SVE; i.e., equilibrium state without external pressure), \Eq{virial} becomes:
\begin{equation}
\alpha_{\rm vir} \equiv \frac{3\sigma^2}{\Gamma GM_{\rm cl}/R} = \frac{5\sigma^2R}{GM_{\rm cl}}\,,
\label{eqn:alpha}
\end{equation}
where we have defined the virial parameter $\alpha_{\rm vir}$.

\AP{i do not fully follow the logic of the following passages}
\DL{need some work to clarify this para}
Motivated by the Larson's relations, we express the observable quantities in the following forms:
\begin{equation}
V_0^2\equiv\frac{\sigma^2}{R},\,
\end{equation}
which is related to Larson's linewidth-size relation \citep{Larson81a}, which can be recast into the form of:
\begin{equation}
\sigma^2/R = \pi G \Sigma/5.
\end{equation}
For pressure-bounded virial equilibrium (PVE; i.e. equilibrium state with pressure), % external pressure
Equation~\ref{eqn:virial} becomes:
\begin{equation}
P_{\rm ext} = \frac{3\sigma^2M_{\rm cl}}{4\pi R^3} - \frac{\Gamma G M_{\rm cl}^2}{4\pi R^4},
\label{eqn:pext}
\end{equation}
which describes the external pressure needed to confine the gas inside a volume $V$. \AP{please specify which eq.s you are referring to} Rearranging the equation and substituting for the surface density $\Sigma$\eq$M/\pi R^2$, we can express this equation using the following:
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \Gamma G \Sigma \pi \right).
\label{eqn:v0}
\end{equation}
Therefore, for a virialized cloud, one expects a one-to-one mapping between $V_0^2$ and $\Sigma$ since $\sigma^2/R\propto\Sigma$, whereas Equation~\ref{eqn:v0} represents the locus along which external pressures $P_{\rm ext}$ are needed in order for MCs to have linewidths $\sigma$ for a given set of surface densities. % (see\Fig{alpha27}).
In other words, \AP{the following is a bit unclear} a cloud with $\sigma$ and $\Sigma$ offset from $\sigma^2/R\mapsto\Sigma$ is in equilibrium with the external pressure $P_{\rm ext}$ (and therefore $V_0^2$ is not constant with $\Sigma$ but varies depending on $P_{\rm ext}$; cf. expectation for SVE, where $\sigma\propto R^{0.5}$; see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}).

Summarizing, to quantify the stability of MC we use the virial parameter (\Eq{alpha}), which describes the balance between kinetic and gravitational potential energies. An $\alpha_{\rm vir}\approx$\,2 corresponds to approximately equipartition between the two energy terms and is often used to assess the boundedness of a given structure in \obs \citep[see e.g., ][]{Kauffmann17b}\footnote{Note, however, that the true virial state/boundedness of a given structure also depends on the surface terms and the magnetic field (see \Eq{virial}).}.
To facilitate comparison with \obs, we derive $\alpha_{\rm vir}$ using the gas component only and ignore contributions due to 
the stellar component.

@@

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}, or the Toomre $Q$ parameter (\AP{i would put this parenthesis as a note later on, i.e. close to \Eq{q_eff}} or the proxy, circular velocity to dispersion ratio; $v_{\rm circ}/\sigma$; e.g., \citealt{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}).
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk of 
finite thickness $h$ is described by 
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,,
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and the third term describes internal pressure support. Then the Toomre-Q parameter can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma\kappa}{\pi G \Sigma}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), the solution of \Eq{3Ddisp} is unstable for scales $k$ such that $Q < 1$ (or equivalently $\omega^2 < 0$). Physically, this situation happens when local gravitational instability 
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen.
% but further evolution depends on their EOS. (e.g., instability impose by angular momentum from gas accretion from the cosmic web and satellite galaxies and stellar feedback).
% Note that fragmentation is expected to happen at critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$, which is greater than the typical size of  GMCs. It should be clear that the instability sets the scales for fragmentation, but star-forming GMCs are the collapsing, denser, and cooler molecular structures on even smaller scales. The fragmented structures thus correspond to regions where $Q\lesssim1.
Similarly, for a collisionless fluid -- as an ensamble of stars -- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\rm star} \equiv\frac{\sigma\kappa}{3.36 G \Sigma}\,.
\end{equation}

In our stability analysis, we account for a non-negligible disk thickness and for the combined effect of gas and stars; this is done adopting the approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modify the $Q$ parameter for gas/star by accounting for the vertical velocity dispersion:
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r 
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
%
The combined effect of gas and stars can then be account by writing
\begin{equation}\label{eqn:q_eff}
%Q^{-1}_{\rm eff} =  \left\{
%				\begin{array}{lccr}
%					     {\displaystyle\frac{w}{T_{\rm star} Q_{\rm star}} + \frac{1}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \geq T_{\rm gas} Q_{\rm gas} \\ [0.75em]
%                                               {\displaystyle\frac{1}{T_{\rm star} Q_{\rm star}} + \frac{w}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \leq T_{\rm gas} Q_{\rm gas} \\				
%				\end{array} 
%			    \right.
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\rm star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\rm star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\rm star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\rm star} \leq Q^{\rm thick}_{\rm gas} \\				
				\end{array} 
			    \right.
\end{equation}
where the relative weight $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\rm star} \sigma_{\rm gas}}{\sigma_{\rm star}^2 + \sigma_{\rm gas}^2},
\end{equation}
\AP{symbol already used for internal enrgy in \Eq{virial_th_general}; do we already need the virial theorem derivation?}
\DL{this is small $w$. Whether or not we need the Larson's relation stuff, I think those eqns there help understand the motivation 
behind relating aa with bb and bb and cc, etc, in the Larson's relations?}

@@

Based on the scaling of $M_J$\eq$\sigma^4/(G^2\Sigma)$ for the Jeans mass, a velocity dispersion of $\sigma\approx50$\,\kms and a gas mass surface density of $\Sigma_{\rm gas}\approx450$\,\Msun\,pc$^{-2}$ yield $M_J\sim$\,8\E{8}\,\Msun, which is in good agreement with the mass range found in the MCs identified (see \Fig{dist}).
%
Similarly massive molecular clouds have been reported in idealized closed-box isolated galaxy simulations done at higher resolution (\AP{which was Behrendt16a resolution? I would report the value} e.g., \citealt{Behrendt16a}). Note, however, that the IGM and merger and accretion histories are not properly modeled in such simulations since they adopt non-cosmological initial conditions. That said, this is reassuring --- our results are not far off in spite of the limited resolution ($l\approx$\,30\,pc).

>>>>

\begin{abstract}
We study the dynamical properties of molecular cloud complexes (MC) in a prototypical galaxy (\flower) 
at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
%
We identify MCs using H$_2$ density, and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$) with those observed in nearby and \z$\sim$\,2 galaxies.
%
The MCs span a mass range of $10^{6.5-9}$\,\Msun and a size range of $R\lesssim200$\,pc. Their velocity dispersions and gas surface densities are systematically higher than those observed in nearby galaxies, but comparable to starburst galaxies.
%
We compute the effective two-component Toomre $Q_{\rm eff}$ parameter (gas and stars) to examine regions that are 
gravitationally unstable.
%
We further perform virial analysis, which is accessible to observations, to examine regions that are likely 
to undergo collapse. Our analysis indicates that MCs in the main disk of \flower are unbound, but its 
substructures have lower virial parameters and Toomre $Q_{\rm eff}$, which is consistent 
with the notion that collapsing structures occur within globally stable structures.
%
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback 
compared to \flower. Our results suggest that the more massive and bigger MCs compared to the Milky Way manifest from the 
higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA) can furnish scrutinizing data to test our findings and shed light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

\begin{figure*}[!htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.65\textwidth]{\figpath/sfh.pdf}
\caption{
    Star formation history of \flower (top) and projected stellar mass distribution (bottom) of \flower during {\it (a)} 
    one of its accretion phases at its early stage of evolution; {\it (b)} 
    one of its major starburst phases after a merger event; and {\it (c)} 
    a relatively quiescent phase post-starburst.
\label{fig:SFH}}
\end{figure*}

@@

In the ``clump-finding'' process, we employ a set of different density thresholds defined based on the molecular hydrogen density of \flower 
at different evolutionary stages (between redshifts of $z$\eq6.0\,$-$\,7.2). We 
note that this process is in essence similar to identifying molecular structures based on the noise levels of surface density maps observers obtain with telescopes, using molecular line tracers such as CO, CS, and HCN, as commonly adopted in observational studies (e.g., identifying ``clumps'' based on/after applying $\sigma$-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a}).

@@

\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples showing the MCs identified in the accreting phase of \flower by applying volumetric H$_2$ density cuts of $n_{\rm cut}$\eq[0.53, 6.81,19.00]\,cm$^{-3}$, which is a subset of all the $n_{\rm cut}$ adopted (see \Sec{method}). Color shows the H$_2$ surface density, weighted by the column density. 
\label{fig:MC}}
% \addtocounter{figure}{-1}
\end{figure*}

In \Fig{h2density}, we show an example of the volumetric H$_2$ density ($n_{\rm H2}$) distribution of \flower~for a given evolutionary stage which includes contributions within $R\sim$\,3.5\,kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of density where ``clumps/structures'' are found based on the Euler characteristic, i.e., the fourth Minkowsky functional (see \citealt{Pallottini17b}). We identify MCs by applying 10 equally-spaced volumetric H$_2$ density cuts of $\log{(n_{\rm cut}/{\rm cm^{-3})}}\in[-0.5, 1.5]$\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 19.00, 31.62]\footnote{We also vary the range of $n_{\rm cut}$ adopted and find no qualitative differences affecting the results and conclusions of this work except in the slope of the CMF (see \Sec{ncut}).} to each snapshot.

For visualization purposes, we overplot the molecular structures identified using a subset of the volumetric H2 density cuts
($n_{\rm cut}$\eq0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps as an example (see \Fig{MC}). 
The plots shown in \Fig{MC} are weighted by the column density of the total gas content.
% equation below for bookkeeping:
% ${\rm pixel_i}$\eq$\frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}$,
% where $\rho_{\rm H2, i}$ is the H$_2$ density of pixel $i$, $w_i$ is the weighted quantity in pixel $i$ which we adopt
% to be the total gas density, and $\Delta z_i$ is the path length of pixel $i$ along the line-of-sight over which the quantities are integrated along.

@@

We show in \Fig{dist} the MC distributions in terms of their molecular gas masses, radii, and gas mass fractions, which we define as
\begin{equation}
f_{\rm gas} = \frac{M_{\rm cl}} {\left(M_{\rm cl} + M_*\right)},
\end{equation}
where $M_*$ is the stellar mass within the MC volume.
%
We note that the distributions vary for different evolutionary stage of \flower. The highest density threshold of $n_{\rm cut}$\eq31.62\,\cc yields a ``minimum'' MC mass of the order of 10$^{6.5}$\,\Msun for the densest MC.
%
We find $\sim$10$-$13 MCs with masses exceeding 10$^8$\,\Msun across all the evolutionary stages considered in this work,
for each of the given $n_{\rm cut}$ (see \Sec{method}). The mass of the MCs identified ranges between $M_{\rm cl}\approx$10$^{6.5-9}$\,\Msun, consistent with those observed in \z$\sim$2 galaxies in rest-frame {\it UV and optical light} \citep{Elmegreen07a, Elmegreen09a}.

@@

% Relating Larson's law to stability and pressure (perhaps move this to other section?).
In general, the virial theorem for a distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot I = 2(T - T_{\rm ext}) + W,
\end{equation}
where $\ddot I$ is the second time derivative of the Lagrangian moment of inertia, $T$ is the internal energy of the gas, $T_s$ is external pressure support, and $W$ is the gravity energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and velocity dispersion $\sigma$, accounting for both thermal and turbulent velocity dispersion. Neglecting magnetic field and calling $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}\ddot I = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R - \frac{\Gamma GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the first term on the right hand side (RHS) is the pressure term from internal velocity dispersion, the second term the external pressure term, and the third term is is the gravity term. For the case of simple virial equilibrium (SVE; i.e., equilibrium state without external pressure), \Eq{virial} becomes:
\begin{equation}
\alpha_{\rm vir} \equiv \frac{3\sigma^2}{\Gamma GM_{\rm cl}/R} = \frac{5\sigma^2R}{GM_{\rm cl}}\,,
\label{eqn:alpha}
\end{equation}
which is also the defined as the virial parameter $\alpha_{\rm vir}$.

Motivated by the Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, 
we recast Equation~\ref{eqn:alpha} into the form of:
\begin{equation}
\sigma^2/R = \pi G \Sigma/5\,, 
\label{eqn:SVE}
\end{equation}
for gas in virial equilibrium (i.e., by setting $\alpha_{\rm vir}$\eq1).
On the other hand, for pressure-bounded virial equilibrium (PVE; i.e. equilibrium state with pressure), % external pressure
Equation~\ref{eqn:virial} becomes:
\begin{equation}
P_{\rm ext} = \frac{3\sigma^2M_{\rm cl}}{4\pi R^3} - \frac{\Gamma G M_{\rm cl}^2}{4\pi R^4},
\label{eqn:pext}
\end{equation}
which describes how much external pressure is needed to confine the gas inside a given volume. 
Substituting surface density $\Sigma$\eq$M/\pi R^2$ into Equation~\ref{eqn:pext}, we can express it with the following:
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \Gamma G \Sigma \pi \right)\,,
\label{eqn:v0}
\end{equation}
keeping $\sigma^2/R$ on the LHS.

Based on Equations~\ref{eqn:SVE} and \ref{eqn:pext}, a one-to-one mapping 
between $\sigma^2/R$ and $\Sigma$ is therefore expected for a virialized cloud, 
since $\sigma^2/R\propto\Sigma$. Equation~\ref{eqn:v0} 
then yields the loci along which external pressures $P_{\rm ext}$ are needed in order for some given MCs to have 
linewidths $\sigma$ for a given set of surface densities. % (see\Fig{alpha27}).
In other words, a cloud with $\sigma$ and $\Sigma$ offset from $\sigma^2/R\mapsto\Sigma$ is in equilibrium with 
external pressure $P_{\rm ext}$ (and therefore $\sigma^2/R$ is not constant 
with $\Sigma$ but varies depending on $P_{\rm ext}$; cf. 
expectation for SVE, where $\sigma\propto R^{0.5}$; see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}).

Summarizing, we use the virial parameter (\Eq{alpha}) to quantify the stability of an MC. 
An $\alpha_{\rm vir}\approx$\,2 corresponds to approximately equipartition between gravitational and kinetic energies, and is
 often used to assess the boundedness of a given structure in \obs \citep[see e.g., ][]{Kauffmann17b}\footnote{Note, however, that the true virial state/boundedness of a given structure also depends on the surface terms and the magnetic field (see \Eq{virial}).}.
To facilitate comparison with \obs, we derive $\alpha_{\rm vir}$ using the gas component only and ignore contributions due to 
the stellar component.

@@

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}, or the Toomre $Q$ parameter (\AP{i would put this parenthesis as a note later on, i.e. close to \Eq{q_eff}} or the proxy, circular velocity to dispersion ratio; $v_{\rm circ}/\sigma$; e.g., \citealt{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}).
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk of 
finite thickness $h$ is described by 
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,,
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and the third term describes internal pressure support. Then the Toomre-Q parameter can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma\kappa}{\pi G \Sigma}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), the solution of \Eq{3Ddisp} is unstable for scales $k$ such that $Q < 1$ (or equivalently $\omega^2 < 0$). Physically, this situation happens when local gravitational instability 
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen.
% but further evolution depends on their EOS. (e.g., instability impose by angular momentum from gas accretion from the cosmic web and satellite galaxies and stellar feedback).
% Note that fragmentation is expected to happen at critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$, which is greater than the typical size of  GMCs. It should be clear that the instability sets the scales for fragmentation, but star-forming GMCs are the collapsing, denser, and cooler molecular structures on even smaller scales. The fragmented structures thus correspond to regions where $Q\lesssim1.
Similarly, for a collisionless fluid -- as an ensamble of stars -- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\star} \equiv\frac{\sigma\kappa}{3.36 G \Sigma}\,.
\end{equation}

In our stability analysis, we account for a non-negligible disk thickness and for the combined effect of gas and stars; this is done adopting the approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modify the $Q$ parameter for gas/star by accounting for the vertical velocity dispersion:
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r 
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
%
The combined effect of gas and stars can then be account by writing
\begin{equation}\label{eqn:q_eff}
%Q^{-1}_{\rm eff} =  \left\{
%				\begin{array}{lccr}
%					     {\displaystyle\frac{w}{T_{\rm star} Q_{\star}} + \frac{1}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\star} \geq T_{\rm gas} Q_{\rm gas} \\ [0.75em]
%                                               {\displaystyle\frac{1}{T_{\rm star} Q_{\star}} + \frac{w}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\star} \leq T_{\rm gas} Q_{\rm gas} \\				
%				\end{array} 
%			    \right.
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\rm star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\rm star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\rm star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\rm star} \leq Q^{\rm thick}_{\rm gas} \\				
				\end{array} 
			    \right.
\end{equation}
where the relative weight $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
\AP{symbol already used for internal enrgy in \Eq{virial_th_general}; do we already need the virial theorem derivation?}
\DL{this is small $w$. Whether or not we need the Larson's relation stuff, I think those eqns there help understand the motivation 
behind relating aa with bb and bb and cc, etc, in the Larson's relations?}

@@

Based on the scaling of $M_J$\eq$\sigma^4/(G^2\Sigma)$ for the Jeans mass, a velocity dispersion of $\sigma\approx50$\,\kms and a gas mass surface density of $\Sigma_{\rm gas}\approx450$\,\Msun\,pc$^{-2}$ yield $M_J\sim$\,8\E{8}\,\Msun, which is in good agreement with the mass range found in the MCs identified (see \Fig{dist}).
%
Similarly massive molecular clouds have been reported in idealized closed-box isolated galaxy simulations done at higher resolution (e.g., a maximum 
resolution of 3\,kpc in the 48\,kpc box studied by \citealt{Behrendt16a}). 
Note, however, that the IGM and merger and accretion histories are not properly modeled in such simulations since they adopt non-cosmological initial conditions. That said, this is reassuring --- our results are not far off in spite of the limited resolution ($l\approx$\,30\,pc).
